---
title: "parameter recovery"
source: "https://stats.stackexchange.com/questions/447029/parameter-recovery-gamma-distribution-and-model-with-a-random-intercept"
output: html_notebook
---

The author was trying to simulate data from a gamma distribution and then fit a Generalized Linear Mixed Model (GLMM) to recover the parameters. The parameter recovery for the fixed effects is agreeable but the random intercept parameter doesn't seem correct.

```{r}
library(tidyverse)
library(lme4)
```

The author was simulating some data from a scenario where there are,

-   40 subjects, who are participating in a longitudinal study with data collected at,

-   3 timepoints (T0, T1, T2) with,

-   30 replicates at each timepoint. In the code below, this is done using a function with a number of different arguments to specify the subjects, repeats, fixed effects values as well as the random intercept SD.

```{r}
subjects <- 40 # number of subjects
repeats  <- 30 # number of replicates per time (3*repeats per subject)
beta_T0   <-  0.2  # Intercept = exp(beta_0)
beta_T1  <-  -0.1  # Effect at Time 1 relative to base-line effect
beta_T2  <-  0.15  # Effect at Time 2 relative to base-line effect 
sigmaI <- 0.2 # between-subject SD
```

disassemble 'GenData' function

```{r}
df_data <- expand.grid(I = 1:subjects, r = 1:repeats, T = 0:2) %>% #expand.grid is equal to multi-for
    arrange(I, T, r)
df_data
```

```{r}
# Random intercept for each individual

df_I <- tibble(I = unique(df_data$I))
df_I$deltaI <- rnorm(nrow(df_I), 0, sigmaI)
df_I
```

```{r}
# Beta intercepts for Time factor

df_beta <- tibble(T = c(0,1,2),
                  beta = c(beta_T0, 
                           beta_T0 + beta_T1, 
                           beta_T0 + beta_T2))
df_beta 
```

```{r}
# Combining it together and taking exponential of the mean

df_data <- df_data %>% 
  left_join(df_I, by = "I") %>% 
  left_join(df_beta, by = "T") %>%
  mutate(
    ln_mu  = beta + deltaI,
    mu     = exp(ln_mu)
    ) 
df_data
```

```{r}
# Calculating the shape parameter using the formula, mean = shape x scale and generating individual values from a gamma distribution

df_data$shape <- df_data$mu / 0.15 
df_data$y <- 0.0
for(i in 1:nrow(df_data)) {
  df_data$y[i] <- rgamma(1,shape = df_data$shape[i], scale = 0.15)
}
df_data
```

```{r}
# Setting as factors

df_data$I   <- factor(df_data$I)
df_data$T   <- factor(df_data$T)
df_data
```

'GenData' function

```{r}
GenData <- function(subjects, repeats, beta_0, beta_T1, beta_T2, sigmaI) {

  # Setting seed

  set.seed(248)

  # Creating data frame

  df_data <- expand.grid(I = 1:subjects, r = 1:repeats, T = 0:2) %>% 
    arrange(I, T, r)

  # Random intercept for each individual

  df_I <- tibble(I = unique(df_data$I))
  df_I$deltaI <- rnorm(nrow(df_I), 0, sigmaI)

  # Beta intercepts for Time factor

  df_beta <- tibble(T = c(0,1,2),
                    beta = c(beta_T0, beta_T0 + beta_T1, beta_T0 + beta_T2))

  # Combining it together and taking exponential of the mean

  df_data <- df_data %>% 
    left_join(df_I, by = "I") %>% 
    left_join(df_beta, by = "T") %>%
    mutate(
      ln_mu  = beta + deltaI,
      mu     = exp(ln_mu)
      ) 

  # Calculating the shape parameter using the formula, mean = shape x scale and generating individual values from a gamma distribution

  df_data$shape <- df_data$mu / 0.15 
  df_data$y <- 0.0
  for(i in 1:nrow(df_data)) {
    df_data$y[i] <- rgamma(1,shape = df_data$shape[i], scale = 0.15)
    }

  # Setting as factors

  df_data$I   <- factor(df_data$I)
  df_data$T   <- factor(df_data$T)

  return (df_data)
}
```

Generating data

```{r}
df_data <- GenData(subjects, repeats, beta_T0, beta_T1, beta_T2, sigmaI)

```

Specifying model and getting model summary

```{r}
fit <- glmer(y ~ 1 + T + (1|I), 
             family = Gamma(link = "log"),
             data = df_data)

summary(fit)
```
